---
title: "Investigate washout"
output: pdf_document
---

Some studies have washout, which can mean different things for different studies.


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
source("read-data.R")
df <- read.data()
```

These are all the studies that are labeled as having washout:


```{r}
kable(df %>% filter(WashOut == 'Y') %>% 
  select(JournalVolumePage, Author, Year, subtype, TypesofMIGSifany))
```

This study has pre- and post-washout and no measurements without washout. This is
equivalent to setting the pre and post op meds to 0:

```{r}
kable(df %>% filter(Washoutbaseline == PreOpIOPMean, !is.na(WashoutIOP)) %>% 
  select(JournalVolumePage, Author, Year, TypesofMIGSifany))
```

This study has preop washout and no washout in the post period. Virgin patients
were enrolled, and had 0 pre-op meds; only after were they put on meds:

```{r}
kable(df %>% filter(Washoutbaseline == PreOpIOPMean, is.na(WashoutIOP)) %>% 
  select(JournalVolumePage, Author, Year, TypesofMIGSifany))
```

These studies had both washout and regular measurements in the pre-period. That 
tells us about the relationship between meds and IOP:

```{r}
kable(df %>% filter(Washoutbaseline != PreOpIOPMean) %>% 
  mutate(mm.Hg.per.med = (Washoutbaseline - PreOpIOPMean)/RxPreOpMean,
         rel.p = 100*(1-(Washoutbaseline - PreOpIOPMean)/Washoutbaseline),
         rel.p.drop.per.med = 
           100*(1-exp(log(rel.p / 100) / RxPreOpMean))) %>%
  select(Author, Year, PreOpIOPMean, RxPreOpMean, mm.Hg.per.med, 
         rel.p, rel.p.drop.per.med), digits = 2)
```

Most commonly, 1 med corresponds to about a 15-20% drop in IOP; or a 4-5 mmHg 
drop. 

This study has washout only in the last period; we can also use this to estimate
the drop in IOP per med:

```{r}
kable(df %>% filter(regexpr("Fea", Author)==TRUE) %>% 
  mutate(mm.Hg.per.med = (WashoutIOP - LastPeriodIOPMean)/RxPostOpMean,
         rel.p = 100*(1-(WashoutIOP - LastPeriodIOPMean)/WashoutIOP),
         rel.p.drop.per.med = 100*(1-exp(log(rel.p / 100) / RxPostOpMean))) %>%
  select(Author, Year, LastPeriodIOPMean, RxPostOpMean, mm.Hg.per.med, 
         rel.p, rel.p.drop.per.med), digits = 2)
```

This study is also in line with the estimate of 4-5mmHg, 15-20% drop in IOP per 
med.

# Loose ends

  * Neiween: the IOP drop per med -- double check these numbers.
  
  * Pfeiffer: I don't understand why the numbers for washout preop don't 
    match with PreOpIOPmean, implying that we have both washout and natural for
    preop period, yet they match in the second washout period. If there's only
    washout in the pre-period, remove the numbers from the second washout period
    column.


