---
title: "Phaco meta analysis"
author: "Patrick Mineault"
date: "14/02/2017"
output: pdf_document
---

# Load data
We load data from a CSV exported from Stata. The `Mo` variables refer to what happens after 6 months. The letter variables `Z`, `AA` etc. refer to what happens after 12 months. That's a bug in how Stata exports names of variables which start with a number - the columns were named `6mo...` and `12mo...`.

```{r} 
library(tidyverse)
library(meta)
library(ggrepel)

source("read-data.R")

setwd("~/Documents/phaco")
df <- read.data()
```

# Summary plot
```{r fig.width = 10, fig.height = 10}
ggplot(df, aes(x = PreOpEyes, y = 
                 ifelse(!is.na(LastPeriodIOPMean) & !is.na(PreOpIOPMean), 
                        LastPeriodIOPMean - PreOpIOPMean,
                        ifelse(!is.na(LastPeriodAbsIOPChangeMean),
                               LastPeriodAbsIOPChangeMean,
                               ifelse(!is.na(OneYAbsIOPChangeMean),
                                      OneYAbsIOPChangeMean,
                                      OneYIOPMean - PreOpIOPMean
                                      )))
                               , label = study.name, color = subtype)) + geom_point() + ylab('IOP change last period') + geom_text_repel()
```


# Analysis without imputation

## 6 month follow-up

```{r fig.width=10, fig.height=5}
m <- metagen(SixMoAbsIOPChangeMean, 
             SixMoAbsIOPChangeStdDev / sqrt(SixMoEyes), 
             study.name, 
             data=df,
             subset=!is.na(df$SixMoEyes))
forest(m, comb.fixed=FALSE)
```

## 12-month follow up

```{r fig.width=10, fig.height=5}
m <- metagen(OneYAbsIOPChangeMean, 
             OneYAbsIOPChangeStdDev / sqrt(OneYEyes), 
             study.name, 
             data=df,
             subset=!is.na(df$OneYAbsIOPChangeMean))
forest(m, comb.fixed = FALSE)
```
## Last period

```{r fig.width=10, fig.height=5}
m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df,
             subset=!is.na(df$LastPeriodAbsIOPChangeMean))
forest(m, comb.fixed = FALSE)
```

## Subgroup analysis

```{r fig.width=10, fig.height=9}
m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df,
             subset=!is.na(df$LastPeriodAbsIOPChangeMean),
             byvar=subtype)
forest(m, comb.fixed = FALSE)
```

There's too much missing data to anything useful without imputation. 

# Analyses with imputation

I'm going to start by imputing on the basis of a low loss in follow up, high correlation between pre and post measures, and no change in relative IOP for those patients that were lost in follow up.

```{r}
# Consider two range of loss of follow up. 
losses <- list(lo=c(.94, .91, .87),
               hi=c(.82, .72, .51))

# Assume a range of correlations between PreOp and PostOp periods.
corrs <- list(lo=c(.25, .25, .25),
              hi=c(.45, .45, .45))

# And a range of outcomes for the missing eyes.
deltas <- list(lo=c(0, 0, 0),  # Same outcome in the unseen arm.
               hi=c(5, 5, 5))  # Worse outcome in the unseen arm.

infer.mean.sem <- function(N.b, m.b, s.b, N.a, m.a, s.a, m.d, s.d, rho, m.delta) {
  # Infer the mean and SEM of the deviation in the metric measured before and after the intervention.
  # N.b, m.b, s.b: N, mean, S.D. of metric before the intervention
  # N.a, m.a, s.a: N, mean, S.D. of metric after the intervention
  # m.d, s.d: mean, S.D. of metric after intervention minus the metric before the intervention for
  #   the group of survivors (N.a). Can be NA.
  # rho: Assumed correlation between before and after scores. Will be used to infer s.d. if s.d is NA
  # m.delta: the assumed mean delta between the m.b for the non-survivors minus the survivors.
  # Use equations in Section 6.1 of Schwarzer, Carpenter & Rucker (2014), Meta-Analysis with R
  m.d <- ifelse(is.na(m.d), m.a - m.b, m.d)
  s.d <- ifelse(is.na(s.d), sqrt(s.a ** 2 + s.b ** 2 - 2 * rho * s.a * s.b), s.d)
  
  # Now generalize the mean difference from the observed subset to the full dataset.
  # pi := unobserved portion
  pi = (N.b - N.a) / N.b
  m.d.full <- m.d + pi * m.delta
  # Let's assume that delta ~ N(m.delta, s.d)
  sem.d.full <- sqrt(1 + pi ** 2) * s.d / sqrt(N.a)
  return(data.frame(m=m.d.full, sem=sem.d.full))
}

library(testthat)
expect_equal(infer.mean.sem(25, 0, 1, 25, 0, 1, NA, NA, 0, 0)$sem, sqrt(2)/sqrt(25))
expect_equal(infer.mean.sem(25, 0, 1, 25, 0, 1, NA, NA, 0.5, 0)$sem, 1/sqrt(25))
expect_equal(infer.mean.sem(25, 0, 1, 25, 0, 1, NA, NA, 0, 0)$m, 0)
expect_equal(infer.mean.sem(25, 0, 1, 16, 0, 1, NA, NA, 0, 0)$sem, sqrt(1 + (9 / 25) ** 2) * sqrt(2)/sqrt(16))
expect_equal(infer.mean.sem(25, -1, 1, 16, 2, 1, NA, NA, 0, 0)$m, 3)
expect_equal(infer.mean.sem(25, -1, 1, 16, 2, 1, NA, NA, 0, 0)$m, 3)
expect_equal(infer.mean.sem(20, 5, 1, 20, -5, 1, NA, NA, 0, 10)$m, -10)
expect_equal(infer.mean.sem(20, 5, 1, 10, -5, 1, NA, NA, 0, 20)$m, 0)

impute.df <- function(df, loss, corr, delta) {
  # Impute missing data using the bone-headed method of just assuming the mean effect of 
  # - loss of follow-up
  # - correlation between Pre and Post IOP metrics
  # - delta between IOP of eyes that were lost in follow up and eyes that were ok. 
  df <- df %>% mutate(imp.SixMoEyes = ifelse(is.na(SixMoEyes), round(loss[1] * PreOpEyes), SixMoEyes),
                      imp.OneYEyes = ifelse(is.na(OneYEyes), round(loss[2] * PreOpEyes), OneYEyes),
                      imp.LastPeriodEyes = ifelse(is.na(LastPeriodEyes), round(loss[3] * PreOpEyes), LastPeriodEyes))
  
  df_ <- with(df, infer.mean.sem(PreOpEyes, PreOpIOPMean, PreOpIOPStdDev,
                                 imp.SixMoEyes, SixMoIOPMean, SixMoIOPStdDev,
                                 SixMoAbsIOPChangeMean, SixMoAbsIOPChangeStdDev, corr[1], delta[1]))
  
  df$imp.SixMoIOPChangeMean <- df_$m
  df$imp.SixMoIOPChangeSEM <- df_$sem
  
  df_ <- with(df, infer.mean.sem(PreOpEyes, PreOpIOPMean, PreOpIOPStdDev,
                                 imp.OneYEyes, OneYIOPMean, OneYIOPStdDev,
                                 OneYAbsIOPChangeMean, OneYAbsIOPChangeStdDev, corr[2], delta[2]))
  df$imp.OneYIOPChangeMean <- df_$m
  df$imp.OneYIOPChangeSEM <- df_$sem
  
  df_ <- with(df, infer.mean.sem(PreOpEyes, PreOpIOPMean, PreOpIOPStdDev,
                                 imp.LastPeriodEyes, LastPeriodIOPMean, LastPeriodIOPStdDev,
                                 LastPeriodAbsIOPChangeMean, LastPeriodAbsIOPChangeStdDev, corr[3], delta[3]))
  df$imp.LastPeriodIOPChangeMean <- df_$m
  df$imp.LastPeriodIOPChangeSEM <- df_$sem
  
  # Patch up NAs for std dev of medications.
  df$imp.RxPreOpStdDev <- ifelse(is.na(df$RxPreOpStdDev), 
                                 .5* df$RxPreOpMean + .2, 
                                 pmax(df$RxPreOpStdDev, .2))

  df$imp.RxPostOpStdDev <- ifelse(is.na(df$RxPostOpStdDev), 
                                 .5* df$RxPostOpMean + .2, 
                                 pmax(df$RxPostOpStdDev, .2))

    
  return(df)
}

# Q: are we dealing properly with loss of follow-up means?
# Verify where available if the before, after measurements, and changes match up.

df <- impute.df(df, losses[['lo']], corrs[['hi']], deltas[['lo']])
```

## 6 month follow-up

```{r fig.width=10, fig.height=10}
m <- metagen(imp.SixMoIOPChangeMean, 
             imp.SixMoIOPChangeSEM, 
             study.name, 
             data=df %>% filter(!is.na(imp.SixMoIOPChangeSEM), subtype != 'MIGS') %>% mutate(subtype = droplevels(subtype)),
             byvar=subtype)
forest(m, comb.fixed=FALSE)
```

## 12-month follow up

```{r fig.width=10, fig.height=10}
m <- metagen(imp.OneYIOPChangeMean, 
             imp.OneYIOPChangeSEM, 
             study.name, 
             data=df %>% filter(!is.na(imp.OneYIOPChangeSEM), subtype != 'MIGS') %>% mutate(subtype = droplevels(subtype)),
             byvar=subtype)
forest(m, comb.fixed=FALSE)
```
## Last period

```{r fig.width=10, fig.height=15}
m <- metagen(imp.LastPeriodIOPChangeMean, 
             imp.LastPeriodIOPChangeSEM, 
             study.name, 
             data=df %>% filter(!is.na(imp.LastPeriodIOPChangeSEM)),
             subset=!is.na(imp.LastPeriodIOPChangeSEM),
             byvar=subtype)
forest(m, comb.fixed=FALSE)
```

## Meds

```{r fig.width=10, fig.height=16}
m <- metacont(imp.LastPeriodEyes,
              RxPostOpMean, 
              imp.RxPostOpStdDev,
              PreOpEyes,
              RxPreOpMean, 
              imp.RxPreOpStdDev,
              study.name, 
              data=df,
              subset=!is.na(imp.RxPreOpStdDev) & !is.na(imp.RxPostOpStdDev),
              byvar=subtype)
forest(m, comb.fixed=FALSE)
```

## Multivariate model to integrate across time


## Correlation between meds and drop in IOP

How is IOP drop related to change in meds? Two hypotheses:

  * Those studies that see the largest IOP drops also have drop in meds, as doctors see that can use the newfound *slack* to decrease the number of meds people take
  * The studies that see the largest IOP drops are those that don't change meds, because dropping meds would also increase IOP

So which is it?

```{r}
df <- df %>% mutate(imp.RxChangeMean = RxPostOpMean - RxPreOpMean,
              imp.RxChangeSEM  = sqrt(imp.RxPostOpStdDev ** 2 + imp.RxPreOpStdDev ** 2) / sqrt(imp.LastPeriodEyes))

ggplot(df, aes(x   =imp.LastPeriodIOPChangeMean, 
               xmin=imp.LastPeriodIOPChangeMean - 1.96*imp.LastPeriodIOPChangeSEM,
               xmax=imp.LastPeriodIOPChangeMean + 1.96*imp.LastPeriodIOPChangeSEM,
               y   = imp.RxChangeMean,
               ymin= imp.RxChangeMean - 1.96*imp.RxChangeSEM,
               ymax= imp.RxChangeMean + 1.96*imp.RxChangeSEM,
               color=subtype
               )) + geom_errorbar() + geom_errorbarh()
```

```{r}
ggplot(df %>% filter(subtype=='OAG' & MIGsYorN == 'N'), aes(x   =imp.LastPeriodIOPChangeMean, 
               xmin=imp.LastPeriodIOPChangeMean - 1.96*imp.LastPeriodIOPChangeSEM,
               xmax=imp.LastPeriodIOPChangeMean + 1.96*imp.LastPeriodIOPChangeSEM,
               y   = imp.RxChangeMean,
               ymin= imp.RxChangeMean - 1.96*imp.RxChangeSEM,
               ymax= imp.RxChangeMean + 1.96*imp.RxChangeSEM,
               label=study.name
               )) + geom_errorbar(alpha=.2) + geom_errorbarh(alpha=.2) + ggtitle('OAG only') + geom_text_repel(segment.alpha = 0)
```

In fact, apart from the Pfeiffer et al. (2015) study, there is an apparent positive correlation between the two effect sizes: studies with larger drops in IOP also tend to see larger drops in Rx.

```{r warning=FALSE}
draw.corr <- function(x.mean, x.sem, y.mean, y.sem) {
  d_ <- data.frame(x = rnorm(n = length(x.mean), mean=x.mean, sd = x.sem),
                  y = rnorm(n = length(x.mean), mean=y.mean, sd = y.sem))
  with(d_ %>% filter(!is.na(x) & !is.na(y)), cor(x, y))
}

cat("Mean +- SE correlation, without Pfeiffer et al\n")
df_ <- df %>% filter(!(study.name %in% c("Pfeiffer et. al (2015)")), subtype=='OAG')
drawn.corrs <- with(df_, replicate(n = 100, 
                                   draw.corr(imp.LastPeriodIOPChangeMean, imp.LastPeriodIOPChangeSEM, imp.RxChangeMean, imp.RxChangeSEM)))
mean(drawn.corrs)
sd(drawn.corrs)

cat("Mean +- SE correlation, with Pfeiffer et al\n")
df_ <- df %>% filter(subtype=='OAG')
drawn.corrs <- with(df_, replicate(n = 100, 
                                  draw.corr(imp.LastPeriodIOPChangeMean, imp.LastPeriodIOPChangeSEM, imp.RxChangeMean, imp.RxChangeSEM)))
mean(drawn.corrs)
sd(drawn.corrs)

```

## Joint inferences

Measure the correlation between different outcomes (IOP at 6 months vs. 12 months).

```{r}
ggplot(df, aes(x   =imp.SixMoIOPChangeMean, 
               xmin=imp.SixMoIOPChangeMean - 1.96*imp.SixMoIOPChangeSEM,
               xmax=imp.SixMoIOPChangeMean + 1.96*imp.SixMoIOPChangeSEM,
               y   = imp.OneYIOPChangeMean,
               ymin= imp.OneYIOPChangeMean - 1.96*imp.OneYIOPChangeSEM,
               ymax= imp.OneYIOPChangeMean + 1.96*imp.OneYIOPChangeSEM,
               label=study.name,
               color=subtype
               )) + geom_errorbar() + geom_errorbarh()
```

It's very clear that six months and 12 months IOP are highly correlated. 

```{r warning=FALSE}
df_ <- df %>% filter(subtype=='OAG')
drawn.corrs <- with(df_, replicate(n = 100, 
                                   draw.corr(imp.SixMoIOPChangeMean, imp.SixMoIOPChangeSEM, imp.OneYIOPChangeMean, imp.OneYIOPChangeSEM)))
cat("Mean +- SE correlation, OAG only\n")
print(mean(drawn.corrs))
print(sd(drawn.corrs))

df_ <- df
drawn.corrs <- with(df_, replicate(n = 100, 
                                   draw.corr(imp.SixMoIOPChangeMean, imp.SixMoIOPChangeSEM, imp.OneYIOPChangeMean, imp.OneYIOPChangeSEM)))
cat("Mean +- SE correlation, All subtypes\n")
print(mean(drawn.corrs))
print(sd(drawn.corrs))
```

Let's repeat the same analysis, this time for 12 months vs. last period

```{r warning=FALSE}
df_ <- df %>% filter(subtype=='OAG')
drawn.corrs <- with(df_, replicate(n = 100, 
                         draw.corr(imp.LastPeriodIOPChangeMean, imp.LastPeriodIOPChangeSEM, imp.OneYIOPChangeMean, imp.OneYIOPChangeSEM)))
cat("Mean +- SE correlation, OAG only\n")
print(mean(drawn.corrs))
print(sd(drawn.corrs))

df_ <- df
drawn.corrs <- with(df_, replicate(n = 100, 
                         draw.corr(imp.LastPeriodIOPChangeMean, imp.LastPeriodIOPChangeSEM, imp.OneYIOPChangeMean, imp.OneYIOPChangeSEM)))
cat("Mean +- SE correlation, All subtypes\n")
print(mean(drawn.corrs))
print(sd(drawn.corrs))
```

That's also really high! Finally, 6 months vs. last period

```{r warning=FALSE}
df_ <- df %>% filter(subtype=='OAG')
drawn.corrs <- with(df_, replicate(n = 100, 
                         draw.corr(imp.LastPeriodIOPChangeMean, imp.LastPeriodIOPChangeSEM, imp.SixMoIOPChangeMean, imp.SixMoIOPChangeSEM)))
cat("Mean +- SE correlation, OAG only\n")
print(mean(drawn.corrs))
print(sd(drawn.corrs))

df_ <- df
drawn.corrs <- with(df_, replicate(n = 100, 
                         draw.corr(imp.LastPeriodIOPChangeMean, imp.LastPeriodIOPChangeSEM, imp.SixMoIOPChangeMean, imp.SixMoIOPChangeSEM)))
cat("Mean +- SE correlation, All subtypes\n")
print(mean(drawn.corrs))
print(sd(drawn.corrs))
```

6 months is also highly correlated with the last period. Let's use mvmeta to infer the effect size for both 6 months and 12 months together.

```{r}
library(mvmeta)

get.correlation.matrices.bi <- function(x, y, assumed.rho) {
  S <- list()
  for(i in 1:length(x)) {
    x[i] <- ifelse(is.na(x[i]), y[i], x[i])
    y[i] <- ifelse(is.na(y[i]), x[i], y[i])
    S[[i]] <- matrix(c(x[i] ** 2, x[i] * y[i]* assumed.rho, 
                       x[i] * y[i] * assumed.rho, y[i] ** 2), ncol=2)
  }
  S
}

df_ <- df %>% filter(!is.na(imp.SixMoIOPChangeSEM) | !is.na(imp.OneYIOPChangeSEM), subtype %in% c('OAG', 'ACG'))
thefit <- mvmeta(cbind(imp.SixMoIOPChangeMean, imp.OneYIOPChangeMean) ~ subtype, 
       S=get.correlation.matrices.bi(df_$imp.SixMoIOPChangeSEM, df_$imp.OneYIOPChangeSEM, .7),
       data=df_,
       method="reml")

summary(thefit)
newdata <- data.frame(subtype=c('OAG', 'ACG'))
pred <- predict(thefit, newdata, se=TRUE)
newdata$SixMoIOPChangeMean <- pred$fit[,1]
newdata$OneYIOPChangeMean <- pred$fit[,2]
newdata$SixMoIOPChangeSEM <- pred$se[,1]
newdata$OneYIOPChangeSEM <- pred$se[,2]

library(reshape2)
nd <- melt(newdata)
nd$period <- substr(nd$variable, 0, 4)
nd$metric <- substr(nd$variable, nchar(as.character(nd$variable)) - 3, nchar(as.character(nd$variable)))
nd
df_ <- dcast(nd, formula = subtype + period ~ metric)
df_$period <- relevel(as.factor(df_$period), ref='SixM')
ggplot(df_, aes(x=period, 
                y=Mean, 
                ymin=Mean - 1.96*eSEM, 
                ymax=Mean + 1.96*eSEM, 
                group=subtype,
                fill=subtype)) + geom_ribbon(alpha=.2) + geom_line(aes(color=subtype))

# TODO(Patrick): repeat for three time periods
```

# Impute under every scenario

Look at what happens depending on how we impute the data.

```{r eval=FALSE}
# TODO(Patrick): Fix this.
results <- list()
for(loss in c('lo', 'hi')) {
  for(corr in c('lo', 'hi')) {
    for(delta in c('lo', 'hi')) {
      df <- impute.df(df, losses[[loss]], corrs[[corr]], deltas[[delta]])
      s <- (metagen(imp.SixMoIOPChangeMean, 
                   imp.SixMoIOPChangeSEM, 
                   study.name, 
                   data=df,
                   subset=!is.na(imp.SixMoIOPChangeSEM) & MIGsYorN == 'N',
                   byvar=subtype))
      row <- data.frame(loss=loss, corr=corr, delta=delta, m=s$TE.random.w, ci.lo=s$upper.random.w, ci.hi=s$lower.random.w, period=6, subtype=c('ACG', 'OAG', 'PXG'))
      results[[length(results) + 1]] <- row
      
      s <- (metagen(imp.OneYIOPChangeMean, 
                   imp.OneYIOPChangeSEM, 
                   study.name, 
                   data=df,
                   subset=!is.na(imp.OneYIOPChangeSEM) & MIGsYorN == 'N',
                   byvar=subtype))
      row <- data.frame(loss=loss, corr=corr, delta=delta, m=s$TE.random.w, ci.lo=s$upper.random.w, ci.hi=s$lower.random.w, period=12, subtype=c('ACG', 'OAG', 'PXG'))
      results[[length(results) + 1]] <- row
      
      s <- (metagen(imp.LastPeriodIOPChangeMean, 
                   imp.LastPeriodIOPChangeSEM, 
                   study.name, 
                   data=df,
                   subset=!is.na(imp.LastPeriodIOPChangeSEM) & MIGsYorN == 'N',
                   byvar=subtype))
      row <- data.frame(loss=loss, corr=corr, delta=delta, m=s$TE.random.w, ci.lo=s$upper.random.w, ci.hi=s$lower.random.w, period=18, subtype=c('ACG', 'OAG', 'PXG'))
      results[[length(results) + 1]] <- row
    }
  }
}

all.df <- do.call(rbind, results)

p <- position_dodge(width=1)
ggplot(all.df, aes(x=period, y = m, ymin=ci.lo, ymax=ci.hi, color=subtype)) + geom_pointrange(position=p)

summary(metagen(imp.LastPeriodIOPChangeMean, 
             imp.LastPeriodIOPChangeSEM, 
             study.name, 
             data=df,
             subset=!is.na(imp.LastPeriodIOPChangeSEM) & MIGsYorN == 'N',
             byvar=subtype))

summary(metagen(imp.LastPeriodIOPChangeMean, 
             imp.LastPeriodIOPChangeSEM, 
             study.name, 
             data=df,
             subset=!is.na(imp.LastPeriodIOPChangeSEM) & MIGsYorN == 'N',
             byvar=subtype))

```

# Some notes on the analysis and the studies
  * None of the studies are randomized, except the EAGLE one -- (2016) Azuara-Bianco et al., Lancet
  * There's no control arm in any of the studies
  * Main outcome is IOP drop
    * The older studies are phaco + glaucoma surgery
    * The new ones are phaco + MIGS - minimally invasive glaucoma surgery

## Slicings to look at
  * MIGS
    * Don't look at MIGS 
  * Type of glaucoma:
    * OAG -> open angle glaucoma  ** ~2-3mm **
    * NTG -> normal tension glaucoma ?
    * ACG -> angle closure glaucoma  ** known to be effective **
    * PXG: pseudo-exfoliation ?

## Dimensions to look at - meta-regression
  * Initial severity (IOP before)
  * Size of study (number of eyes)
  * Year
  
## Different outcomes
  * Primary is IOP drop
    * time points 6 mo, 12 mo, (last time point)  
    * most important is 12 months
  * Number of meds
    * Huge confound, because it's controlled by the doctor
    * Meds themselves decrease the IOP
    * A handful of studies use washout pre and post (measuring the IOP without meds) to undo the confounding
        * EAGLE, Samuelson studies have washout
        * Lack of washout will have a tendency to decrease the apparent effectiveness of the studies
    * One med ~= 20% decrease in IOP
    * One med := decrease in quality of life
    * RxPostOpMean is at the same time as LastPeriod
  * (visual acuity but it's kind of obvious)
  
## Additional analyses to perform

  * Funnel plot for small / medium large studies
  * Deal appropriately with multiple arms of same study, e.g. Damji et al., Merz...
  * Deal with three forms of lossiness:
    * Absolutes reported, relatives needed
      * Can patch up using estimate of rho -- sqrt(s_1 ** 2 + s_2 ** 2 - 2 * rho * s_1 * s_2)
      * Try rho = 0, rho = 0.5
    * Loss of follow-up
      * Can deal with by assuming that follow up is either MCAR or worse than MCAR
      * Try mean_delta = 0, mean_delta = -3, mean_delta = -5
    * Not all metrics reported for every study
      * Use mvmeta