---
title: "Phaco meta analysis"
author: "Patrick Mineault"
date: "14/02/2017"
header-includes:
   - \usepackage{bbm}
output: 
  pdf_document:
    latex_engine: xelatex
    toc: yes
---

```{r echo=FALSE, include=FALSE, message=FALSE} 
library(grid)
library(tidyverse)
library(meta)
library(ggrepel)

source("read-data.R")

setwd("~/Documents/phaco")
df <- read.data()
df <- filter.data(df)
```

# Analysis of full dataset

## Main analysis: 12-month+ follow up

```{r fig.width=10, fig.height=10, dev="cairo_pdf"}
df_ <- df %>% 
  filter(!is.na(LastPeriodAbsIOPChangeMean), subtype != "AACG", MIGsYorN == 'N') %>% 
  mutate(subtype=factor(subtype))
m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=LastPeriodEyes)
forest(m, 
       comb.fixed=FALSE, 
       digits=1, 
       digits.se = 2, 
       overall=FALSE, 
       leftcols=c("studlab", "TE", "seTE", "n.e"),
       leftlabs=c("Study", "ΔIOP", "SE", "eyes"))
```

## AACG

```{r results="hold"}
df_ <- df %>% 
  filter(!is.na(LastPeriodAbsIOPChangeMean), subtype == "AACG", MIGsYorN == 'N') %>% 
  mutate(subtype=factor(subtype)) %>% dplyr::arrange(Year)
m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes),
             study.name, 
             data=df_,
             n.e=LastPeriodEyes, comb.fixed = FALSE)
print(m)

df_ <- df %>% 
  filter(!is.na(LastPeriodAbsIOPChangeMean), subtype == "AACG", MIGsYorN == 'N') %>% 
  mutate(subtype=factor(subtype)) %>% dplyr::arrange(Year)
m <- metagen(RxChangeMean,
             RxChangeStdDev / sqrt(LastPeriodEyes),
             study.name, 
             data=df_,
             n.e=LastPeriodEyes, comb.fixed = FALSE)
print(m)

df_ %>% dplyr::select(RxPreOpMean, RxPostOpMean)
```

## Meds

```{r fig.width=10, fig.height=10, dev="cairo_pdf"}
df_ <- df %>% 
  filter(!is.na(LastPeriodAbsIOPChangeMean),
         df$subtype != "AACG", MIGsYorN == 'N') %>% 
  mutate(subtype=factor(subtype))
m <- metagen(RxChangeMean,
             RxChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=LastPeriodEyes)
forest(m, 
       comb.fixed=FALSE, 
       digits=1, 
       digits.se = 2, 
       overall=FALSE, 
       leftcols=c("studlab", "TE", "seTE", "n.e"),
       leftlabs=c("Study", "ΔIOP", "SE", "eyes"))
```



# Correlation between meds and drop in IOP

How is IOP drop related to change in meds? Two hypotheses:

  * Those studies that see the largest IOP drops also have drop in meds, as doctors see that can use the newfound *slack* to decrease the number of meds people take
  * The studies that see the largest IOP drops are those that don't change meds, because dropping meds would also increase IOP

So which is it?

```{r}
df_ <- df %>% mutate(RxChangeSEM  = RxChangeStdDev / sqrt(LastPeriodEyes),
                     LastPeriodAbsIOPChangeSEM = LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes))

ggplot(df_, aes(x  =LastPeriodAbsIOPChangeMean, 
               xmin=LastPeriodAbsIOPChangeMean - 1.96*LastPeriodAbsIOPChangeSEM,
               xmax=LastPeriodAbsIOPChangeMean + 1.96*LastPeriodAbsIOPChangeSEM,
               y   =RxChangeMean,
               ymin=RxChangeMean - 1.96*RxChangeSEM,
               ymax=RxChangeMean + 1.96*RxChangeSEM,
               color=subtype
               )) + geom_errorbar() + geom_errorbarh()
```

```{r warning=FALSE}
ggplot(df_ %>% filter(subtype=='OAG' & MIGsYorN == 'N'), 
        aes(x  =LastPeriodAbsIOPChangeMean, 
            xmin=LastPeriodAbsIOPChangeMean - 1.96*LastPeriodAbsIOPChangeSEM,
            xmax=LastPeriodAbsIOPChangeMean + 1.96*LastPeriodAbsIOPChangeSEM,
            y   =RxChangeMean,
            ymin=RxChangeMean - 1.96*RxChangeSEM,
            ymax=RxChangeMean + 1.96*RxChangeSEM,
            label=study.name
           )) + geom_errorbar(alpha=.2) + geom_errorbarh(alpha=.2) + ggtitle('OAG only') + geom_text_repel(segment.alpha = 0)
```

There is an apparent positive correlation between the two effect sizes: studies with larger drops in IOP also tend to see larger drops in Rx. 

```{r warning=FALSE}
draw.corr <- function(x.mean, x.sem, y.mean, y.sem) {
  d_ <- data.frame(x = rnorm(n = length(x.mean), mean=x.mean, sd = x.sem),
                  y = rnorm(n = length(x.mean), mean=y.mean, sd = y.sem))
  with(d_ %>% filter(!is.na(x) & !is.na(y)), cor(x, y))
}

cat("Mean +- SE correlation, all studies\n")
df_ <- df %>% filter(subtype=='OAG', MIGsYorN == 'N')
drawn.corrs <- with(df_, replicate(n = 100, 
                                   draw.corr(LastPeriodAbsIOPChangeMean, 
                                             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
                                             RxChangeMean, 
                                             RxChangeStdDev / sqrt(LastPeriodEyes))))
mean(drawn.corrs)
sd(drawn.corrs)

cat("Mean +- SE correlation, no washout\n")
df_ <- filter.data(df, 'nowashout') %>% filter(subtype=='OAG', MIGsYorN == 'N')
drawn.corrs <- with(df_, replicate(n = 100, 
                                  draw.corr(LastPeriodAbsIOPChangeMean, 
                                             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
                                             RxChangeMean, 
                                             RxChangeStdDev / sqrt(LastPeriodEyes))))
mean(drawn.corrs)
sd(drawn.corrs)

```

However, this effect goes away when we focus on the studies which don't have 
washout.

# Separate meta-analysis for each time period

```{r message=FALSE}
df <- read.data(fill.last = FALSE)
```

## 6 month follow-up

```{r fig.width=10, fig.height=8, dev="cairo_pdf"}
df_ <- df %>% 
  filter(!is.na(SixMoAbsIOPChangeMean), subtype != "AACG") %>% 
  mutate(subtype=factor(subtype))
m <- metagen(SixMoAbsIOPChangeMean, 
             SixMoAbsIOPChangeStdDev / sqrt(SixMoEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=SixMoEyes)
forest(m, 
       comb.fixed=FALSE, 
       digits=1, 
       digits.se = 2, 
       overall=FALSE, 
       leftcols=c("studlab", "TE", "seTE", "n.e"),
       leftlabs=c("Study", "ΔIOP", "SEM", "eyes"))
```

## 12-month follow up

```{r fig.width=10, fig.height=9, dev="cairo_pdf"}
df_ <- df %>% 
  filter(!is.na(OneYAbsIOPChangeMean), subtype != "AACG", MIGsYorN == 'N') %>% 
  mutate(subtype=factor(subtype))
m <- metagen(OneYAbsIOPChangeMean, 
             OneYAbsIOPChangeStdDev / sqrt(OneYEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=OneYEyes)
forest(m, 
       comb.fixed=FALSE, 
       digits=1, 
       digits.se = 2, 
       overall=FALSE, 
       leftcols=c("studlab", "TE", "seTE", "n.e"),
       leftlabs=c("Study", "ΔIOP", "SEM", "eyes"))
```

## Last period

```{r fig.width=10, fig.height=9, dev="cairo_pdf"}
df_ <- df %>% 
  filter(!is.na(LastPeriodAbsIOPChangeMean), subtype != "AACG", MIGsYorN == 'N') %>% 
  mutate(subtype=factor(subtype))
m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=LastPeriodEyes)
forest(m, 
       comb.fixed=FALSE, 
       digits=1, 
       digits.se = 2, 
       overall=FALSE, 
       leftcols=c("studlab", "TE", "seTE", "n.e"),
       leftlabs=c("Study", "ΔIOP", "SEM", "eyes"))
```

## Correlation among time points

Measure the correlation between different outcomes (IOP at 6 months vs. 12 months).

```{r dev="cairo_pdf"}
ggplot(df %>% filter(subtype != 'AACG') %>% mutate(subtype = factor(subtype)), 
       aes(x   =SixMoAbsIOPChangeMean, 
               xmin=SixMoAbsIOPChangeMean - 1.96*SixMoAbsIOPChangeStdDev / sqrt(SixMoEyes),
               xmax=SixMoAbsIOPChangeMean + 1.96*SixMoAbsIOPChangeStdDev / sqrt(SixMoEyes),
               y   = OneYAbsIOPChangeMean,
               ymin= OneYAbsIOPChangeMean - 1.96*OneYAbsIOPChangeStdDev / sqrt(OneYEyes),
               ymax= OneYAbsIOPChangeMean + 1.96*OneYAbsIOPChangeStdDev / sqrt(OneYEyes),
               label=study.name,
               color=subtype
               )) +
  geom_abline(slope = 1, color="gray70") + 
  geom_errorbar() + 
  geom_errorbarh() + 
  xlab('ΔIOP at 6 months') +
  ylab('ΔIOP at one year') +
  coord_fixed(xlim=c(-18, 0), ylim=c(-18,0)) 
```

It's very clear that six months and 12 months IOP are highly correlated. 

Similarly for one-year vs. last period:

```{r dev="cairo_pdf"}
ggplot(df %>% filter(subtype != 'AACG') %>% mutate(subtype = factor(subtype)), 
       aes(y   =LastPeriodAbsIOPChangeMean, 
               ymin=LastPeriodAbsIOPChangeMean - 1.96*LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes),
               ymax=LastPeriodAbsIOPChangeMean + 1.96*LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes),
               x   = OneYAbsIOPChangeMean,
               xmin= OneYAbsIOPChangeMean - 1.96*OneYAbsIOPChangeStdDev / sqrt(OneYEyes),
               xmax= OneYAbsIOPChangeMean + 1.96*OneYAbsIOPChangeStdDev / sqrt(OneYEyes),
               label=study.name,
               color=subtype
               )) +
  geom_abline(slope = 1, color="gray70") + 
  geom_errorbar() + geom_errorbarh() + 
  coord_fixed(xlim=c(-20, 0), ylim=c(-20,0))
```

Again, correlations are very high. Present this in another way.

```{r fig.height=7, fig.width=5}
library(reshape2)
nd <- melt(df %>% 
             filter(MIGsYorN == 'N',
                    subtype != 'AACG',
                    1*is.na(SixMoAbsIOPChangeMean) + 
                    1*is.na(OneYAbsIOPChangeMean) + 
                    1*is.na(LastPeriodAbsIOPChangeMean) < 2) %>%
             mutate(SixMoChangeSEM = SixMoAbsIOPChangeStdDev / sqrt(SixMoEyes),
                    OneYChangeSEM = OneYAbsIOPChangeStdDev / sqrt(OneYEyes),
                    LastPeriodChangeSEM = LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes)) %>%
             select(study.name, subtype,
                    SixMoAbsIOPChangeMean,
                    OneYAbsIOPChangeMean,
                    LastPeriodAbsIOPChangeMean,
                    SixMoChangeSEM,
                    OneYChangeSEM,
                    LastPeriodChangeSEM), id.vars=c("study.name", "subtype"))
nd$metric <- substr(nd$variable, nchar(as.character(nd$variable)) - 3, nchar(as.character(nd$variable)))
nd$period <- substr(nd$variable, 0, 4)
df_ <- dcast(nd, formula = study.name + subtype + period ~ metric)
df_ <- df_ %>% mutate(period = factor(period, c('SixM', 'OneY', 'Last')),
                      g = paste(study.name, as.character(subtype))) %>% filter(!is.na(Mean))
ggplot(df_, aes(y  =Mean, 
               ymin=Mean - 1.96*eSEM,
               ymax=Mean + 1.96*eSEM,
               x   = period,
               label=study.name,
               group=g)) +
  geom_ribbon(alpha=.02) + 
  geom_line(alpha=.5, aes(color=subtype)) + 
  coord_cartesian(y=c(-15, 0)) + ggtitle('Time course of IOP per study') + theme(plot.title = element_text(hjust = 0.5))

```

It's remarkable how consistent measurements are between time periods. At most, 
we find a change of +- 2.5 mm Hg between the first and last period. 

```{r warning=FALSE}
df_ <- df %>% filter(subtype=='OAG')
drawn.corrs <- with(df_, replicate(n = 100, 
                                   draw.corr(SixMoAbsIOPChangeMean, 
                                             SixMoAbsIOPChangeStdDev / sqrt(SixMoEyes), 
                                             OneYAbsIOPChangeMean, 
                                             OneYAbsIOPChangeStdDev / sqrt(OneYEyes))))
cat("Mean +- SE correlation, OAG only\n")
print(mean(drawn.corrs))
print(sd(drawn.corrs))

df_ <- df
drawn.corrs <- with(df_, replicate(n = 100, 
                                   draw.corr(SixMoAbsIOPChangeMean, 
                                             SixMoAbsIOPChangeStdDev / sqrt(SixMoEyes), 
                                             OneYAbsIOPChangeMean, 
                                             OneYAbsIOPChangeStdDev / sqrt(OneYEyes))))
cat("Mean +- SE correlation, All subtypes\n")
print(mean(drawn.corrs))
print(sd(drawn.corrs))

cat("Regression of one year against 6 months")
print(summary(lm(OneYAbsIOPChangeMean ~ SixMoAbsIOPChangeMean, 
                 df, 
                 weights = OneYAbsIOPChangeStdDev / sqrt(OneYEyes))))
```

This is also reflected in the correlations - even accounting for noise, the correlations
are > .9 between month 6 and month 12. 

# Multivariate inference

Let's use `mvmeta` to infer the effect size for all periods together.

```{r}
library(mvmeta)

fill.na <- function(x, y, z) {
  return(ifelse(!is.na(x), 
                   x,
                   ifelse(is.na(y), 
                          z,
                          ifelse(is.na(z),
                            y,
                            sqrt((y**2 + z**2) / 2 )))))
}

get.correlation.matrices.tri <- function(x, y, z, assumed.rho) {
  S <- list()
  for(i in 1:length(x)) {
    xx <- fill.na(x[i], y[i], z[i])
    yy <- fill.na(y[i], x[i], z[i])
    zz <- fill.na(z[i], x[i], y[i])
    S[[i]] <- matrix(c(xx ** 2, xx * yy * assumed.rho, xx * zz * assumed.rho ** 2,
                       xx * yy * assumed.rho, yy ** 2, zz * yy* assumed.rho,
                       xx * zz * assumed.rho ** 2, zz * yy * assumed.rho, zz * zz), ncol=3)
  }
  S
}

df_ <- df %>% filter(!is.na(LastPeriodAbsIOPChangeStdDev) | 
                     !is.na(SixMoAbsIOPChangeStdDev) | 
                     !is.na(OneYAbsIOPChangeStdDev), subtype %in% c('OAG', 'ACG'), MIGsYorN == 'N')

thefit <- mvmeta(cbind(SixMoAbsIOPChangeMean, OneYAbsIOPChangeMean, LastPeriodAbsIOPChangeMean) ~ subtype, 
       S=get.correlation.matrices.tri(df_$SixMoAbsIOPChangeStdDev / sqrt(df_$SixMoEyes), 
                                      df_$OneYAbsIOPChangeStdDev / sqrt(df_$OneYEyes), 
                                      df_$LastPeriodAbsIOPChangeStdDev / sqrt(df_$LastPeriodEyes), .7),
       data=df_,
       method="reml")

summary(thefit)
newdata <- data.frame(subtype=c('OAG', 'ACG'))
pred <- predict(thefit, newdata, se=TRUE)
newdata$SixMoAbsIOPChangeMean <- pred$fit[,1]
newdata$OneYAbsIOPChangeMean <- pred$fit[,2]
newdata$LastPeriodAbsIOPChangeMean <- pred$fit[,3]
newdata$SixMoAbsIOPChangeSEM <- pred$se[,1]
newdata$OneYAbsIOPChangeSEM <- pred$se[,2]
newdata$LastPeriodAbsIOPChangeSEM <- pred$se[,3]

library(reshape2)
nd <- melt(newdata)
nd$period <- substr(nd$variable, 0, 4)
nd$metric <- substr(nd$variable, nchar(as.character(nd$variable)) - 3, nchar(as.character(nd$variable)))
df_ <- dcast(nd, formula = subtype + period ~ metric)
df_$period <- factor(df_$period, c('SixM', 'OneY', 'Last'))
ggplot(df_, aes(x=period, 
                y=Mean, 
                ymin=Mean - 1.96*eSEM, 
                ymax=Mean + 1.96*eSEM, 
                group=subtype,
                fill=subtype)) + geom_ribbon(alpha=.2) + geom_line(aes(color=subtype)) + ylab("Inferred IOP change per period")
```

# Meta-regression

Consider relationships between different covariates and outcomes. Focus on the 
IOP drop at one year and its correlation with different factors.

```{r}
df <- read.data()
df_ <- df %>% filter(!is.na(LastPeriodAbsIOPChangeMean), 
                     subtype != "AACG", 
                     MIGsYorN == 'N') %>% 
  mutate(coarseWashoutType=factor(washout.type, c("None", "Partial", "Pre", "Both")),
         subtype=relevel(factor(subtype), ref="OAG"))
levels(df_$coarseWashoutType) <- c("None", "None", "Pre", "Both")

m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=OneYEyes)

# Leveled against OAG
print(metareg(~ LastPeriodEyes, x=m))
print(metareg(~ Year, x=m))
print(metareg(~ PreOpIOPMean * subtype + coarseWashoutType + AgeMean + Male, x=m))

# Leveled against ACG
df_ <- df_ %>% mutate(subtype=relevel(factor(subtype), ref="ACG"))

m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=OneYEyes)

print(metareg(~ LastPeriodEyes, x=m))
print(metareg(~ Year, x=m))
print(metareg(~ PreOpIOPMean * subtype + coarseWashoutType + AgeMean + Male, x=m))


# Leveled against PXG
df_ <- df_ %>% mutate(subtype=relevel(factor(subtype), ref="PXG"))

m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=OneYEyes)

print(metareg(~ LastPeriodEyes, x=m))
print(metareg(~ Year, x=m))
print(metareg(~ PreOpIOPMean * subtype + coarseWashoutType + AgeMean + Male, x=m))


# Restricted to OAG only
df %>% 
  filter(subtype == 'OAG', washout.type != 'Both', !is.na(LastPeriodAbsIOPChangeStdDev)) %>% 
  summarize(min = min(PreOpIOPMean),
            max = max(PreOpIOPMean),
            sd = sd(PreOpIOPMean),
            mean = mean(PreOpIOPMean))
```

# Small study bias

```{r}
df_ <- df %>% filter(!is.na(df$LastPeriodAbsIOPChangeMean), 
                     df$subtype != "AACG", 
                     MIGsYorN == 'N') %>% mutate(subtype=factor(subtype))
for(l in levels(df_$subtype)) {
  m <- metagen(LastPeriodAbsIOPChangeMean, 
               LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
               study.name, 
               data=df_ %>% filter(subtype == l),
               n.e=LastPeriodEyes)
  funnel(m)
  title(paste('Funnel plot', l))
}

```

# Alternative filterings of the data

# Prospective studies only

```{r fig.width=10, fig.height=9, echo=FALSE, message=FALSE, dev="cairo_pdf"}
df <- read.data()
df <- filter.data(df, 'prospective')
df_ <- df %>% 
  filter(!is.na(LastPeriodAbsIOPChangeMean), subtype != "AACG", MIGsYorN == 'N') %>% 
  mutate(subtype=factor(subtype))
m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=LastPeriodEyes)
forest(m, 
       comb.fixed=FALSE, 
       digits=1, 
       digits.se = 2, 
       overall=FALSE, 
       leftcols=c("studlab", "TE", "seTE", "n.e"))
```

## Meds

```{r fig.width=10, fig.height=10, dev="cairo_pdf"}
df_ <- df %>% 
  filter(!is.na(RxChangeMean), !is.na(RxChangeStdDev),
         df$subtype != "AACG", MIGsYorN == 'N') %>% 
  mutate(subtype=factor(subtype))
m <- metagen(RxChangeMean,
             sqrt(RxPostOpStdDev** 2 + RxPreOpStdDev ** 2) / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=LastPeriodEyes)
forest(m, 
       comb.fixed=FALSE, 
       digits=1, 
       digits.se = 2, 
       overall=FALSE, 
       leftcols=c("studlab", "TE", "seTE", "n.e"))
```

# Excluding washout studies

```{r fig.width=10, fig.height=9, echo=FALSE, message=FALSE}
df <- read.data()
df <- filter.data(df, 'nowashout')
```

## Last period

```{r fig.width=10, fig.height=9, dev="cairo_pdf"}
df_ <- df %>% 
  filter(!is.na(LastPeriodAbsIOPChangeMean), subtype != "AACG", MIGsYorN == 'N') %>% 
  mutate(subtype=factor(subtype))
m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=LastPeriodEyes)
forest(m, 
       comb.fixed=FALSE, 
       digits=1, 
       digits.se = 2, 
       overall=FALSE, 
       leftcols=c("studlab", "TE", "seTE", "n.e"))
```

## Meds

```{r fig.width=10, fig.height=10}
df_ <- df %>% 
  filter(!is.na(RxChangeMean), !is.na(RxChangeStdDev),
         df$subtype != "AACG", MIGsYorN == 'N') %>% 
  mutate(subtype=factor(subtype))
m <- metagen(RxChangeMean,
             sqrt(RxPostOpStdDev** 2 + RxPreOpStdDev ** 2) / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             byvar=subtype,
             n.e=LastPeriodEyes)
forest(m, 
       comb.fixed=FALSE, 
       digits=1, 
       digits.se = 2, 
       overall=FALSE, 
       leftcols=c("studlab", "TE", "seTE", "n.e"))
```

# Sensitivity to missingness

Simulate what the results would look like if there was no effect in the eyes 
lost to follow up ($\Delta IOP = 0$).

```{r fig.width=10, fig.height=8, dev="cairo_pdf", message=FALSE}
meta.analysis.with.sensitivity <- function(missingness='zero') {
  df <- read.data()
  df <- filter.data(df, 'prospective')
  
  df_ <- df %>% 
    filter(!is.na(LastPeriodAbsIOPChangeMean), subtype != "AACG") %>% 
    mutate(subtype=factor(subtype))
  
  # Simulate a 0 effect in the unobserved fraction.
  df.missing <- df_
  if(missingness == 'zero') {
    # Zero out.
    df.missing <- df.missing %>% mutate(LastPeriodEyes = PreOpEyes - LastPeriodEyes,
                          LastPeriodAbsIOPChangeMean = 0)
  } else {
    # Add 5 mm Hg to each missing eye.
    df.missing <- df.missing %>% mutate(LastPeriodEyes = PreOpEyes - LastPeriodEyes,
                          LastPeriodAbsIOPChangeMean = LastPeriodAbsIOPChangeMean + 5)
  }
  
  df_ <- rbind(df_, df.missing)
  # Aggregate two by two
  for(i in seq(nrow(df.missing), 1)) {
    idx <- rep(FALSE, nrow(df_))
    idx[i] <- TRUE
    idx[i*2] <- TRUE
    df_ <- agg.arms(df_, idx)
  }
  
  df_ <- df_ %>% dplyr::arrange(Year, study.name)
  
  m <- metagen(LastPeriodAbsIOPChangeMean, 
               LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
               study.name, 
               data=df_,
               byvar=subtype,
               n.e=LastPeriodEyes)
  forest(m, 
         comb.fixed=FALSE, 
         digits=1, 
         digits.se = 2, 
         overall=FALSE, 
         leftcols=c("studlab", "TE", "seTE", "n.e"),
         leftlabs=c("Study", "ΔIOP", "SE", "eyes"))
}

meta.analysis.with.sensitivity()
grid.text(paste0("Simulated net change IOP when ΔIOP = 0 in eyes lost to follow up"), .5, .97, gp=gpar(cex=1.3))
```

```{r fig.width=10, fig.height=8, dev="cairo_pdf", message=FALSE}
meta.analysis.with.sensitivity('five')
grid.text(paste0("Simulated net change IOP when ΔIOP = 5 mm Hg higher in eyes lost to follow up"), .5, .97, gp=gpar(cex=1.3))
```

# MIGS

We don't report these results because it's a bit misleading -- the studies
aren't very similar to each other, and we don't use the information in their
control arms. We can do a much better job through, for example, network
meta-analysis, which we plan to do in a future paper.

```{r fig.width=10.5, fig.height=8, dev="cairo_pdf"}
df <- read.data(drop.migs = FALSE)
df_ <- df %>% 
  filter(!is.na(LastPeriodAbsIOPChangeMean), subtype != "AACG", MIGsYorN == 'Y') %>% 
  mutate(subtype=factor(subtype)) %>% dplyr::arrange(TypesofMIGSifany, Year)
m <- metagen(LastPeriodAbsIOPChangeMean, 
             LastPeriodAbsIOPChangeStdDev / sqrt(LastPeriodEyes), 
             study.name, 
             data=df_,
             n.e=LastPeriodEyes)
forest(m, 
       comb.fixed=FALSE, 
       digits=1, 
       digits.se = 2, 
       overall=TRUE, 
       leftcols=c("studlab", "TE", "seTE", "n.e", "TypesofMIGSifany"),
       leftlabs=c("Study", "TE", "seTE", "Total", "Type"))
```