---
title: "Methods and results"
output:
  pdf_document: 
    latex_engine: xelatex
---

> TODO(Patrick): think more deeply about regression to the mean and this will 
affect our claims. Reference:
https://www.reviewofoptometry.com/ce/considerations-in-iop-measurement

Follow the procedures in [PRISMA
2009](http://prisma-statement.org/documents/PRISMA%202009%20checklist.pdf).

# 10. Data extraction

An undergraduate assistant manually extracted data from the tables of the
identified studies. In one case where the data of interest was not cited in the
main study (ref Liu et al.), the data was extracted from figures using a
graphical data extraction tool (WebPlotDigitizer,
http://arohatgi.info/WebPlotDigitizer/; ref Cochrane Review).

The data tables were spot-checked by M.M. against the primary studies. Data was
also audited for self-consistency by P.M. -- e.g. ensuring that differences in
IOP post-surgery were consistent with pre- and post-surgery IOP measurements, 
that number of eyes tracked did not increase through follow-up, etc. The full
extent of these automated checks (unit tests) can be viewed at (ref github 
link).

The data for OAG studies was also cross-verified against Table 1 of Thomas et al
(ref). We rejected one study from meta-analysis because the SD numbers could not
be relied on (reported range too large for reported SD; ref Jacobi et al.). The
full extracted data can be accessed at (ref Github link).

# 11 & 13. Data items & summary measures

We extracted the following columns from the primary sources:

```{r echo=FALSE, warning=FALSE, message=FALSE}
source("read-data.R")
library(knitr)
source('read-data.R')
df <- read.data()
```


```{r message=FALSE}
extracted.data <- read.column.names()
kable(data.frame(extracted.data))
```

> TODO(Marisse): transform this to a table, group similar measurements together.

The main metric of interest for our meta-analysis was the IOP change at the 
final follow-up period (12 months or longer). In complementary analyses, we 
performed meta-analyses separately for 6 month, 12 month, and > 12 month 
follow-ups; the results differed by less than 1 mm per period. and in a 
complementary multivariate meta-analysis we found even these small differences 
shrinked further when taking into account missing follow-up and serial 
correlation.

> TODO(Marisse): decide whether you want to include the information in that last
paragraph in the methods or in the results.

Some studies used washout in at least some time periods to measure the effect of
surgery separately from the effect of medication. When both pre- and
post-surgery IOP were reported with washout, these are the values we used (refs
Pfeiffer, Vold); this is indicated by an asterisk in the main results. When
washout was used in only some periods, we used non-washouted values across time
periods, for consistency. The only exception was (ref Azuara−Blanco 2016), where
IOP was measured with washout in the pre-period but not in the post period, due
to study design. We indicate this study with a double asterisk.

Many studies did not report IOP drop directly, but rather only absolute IOP in
the pre-period and at the follow-up period. Assuming full follow-up, the IOP
drop is simply the difference of these measurements, while the SD ($\sigma$) of
the IOP drop is estimated by:

$\sigma_{drop}^2 = \sigma_{pre}^2 + \sigma_{post}^2 - 2 r \sigma_{pre} \sigma_{post}$

$r$, the correlation between the pre-and-post measurements, was set to the 
median of the correlations in the studies reporting the full set of metrics 
(r = 0.35).

```{r message=FALSE}
df <- read.data(impute.change = FALSE)
rs <- with(df, (PreOpIOPStdDev ** 2 + LastPeriodIOPStdDev ** 2 - LastPeriodAbsIOPChangeStdDev ** 2) / (PreOpIOPStdDev * LastPeriodIOPStdDev)/2)
quantile(rs, .5, na.rm=TRUE)
```
  
Many studies reported several arms corresponding to different severities, 
treatments and subtypes. For phacoemulsification studies, we aggregated 
the data to obtain one arm per glaucoma subtype (ACG, OAG, PXG, acute) 
per study. 

After this step, the total number of arms and eyes per subtype was as follows:

```{r}
df_ <- df %>% filter(!is.na(LastPeriodAbsIOPChangeMean)) %>% group_by(subtype) %>% 
  dplyr::summarize(
    n.total = n(),
    n.prospective = sum(StudyType == 'Prospective'),
    n.retrospective = sum(StudyType == 'Retrospective'),
    PreOpEyes = sum(PreOpEyes),
    FinalPeriodEyes = sum(LastPeriodEyes))

df_ <- df_ %>% rbind(., summarize(df_, 
                                  subtype="(all)",
                                  n.total=sum(n.total),
                                  n.prospective=sum(n.prospective),
                                  n.retrospective=sum(n.retrospective),
                                  PreOpEyes=sum(PreOpEyes),
                                  FinalPeriodEyes = sum(FinalPeriodEyes)))
  
kable(df_)
```

The full list of studies is as follow:

```{r}
df_ <- df %>% filter(!is.na(LastPeriodAbsIOPChangeMean)) %>% 
  arrange(subtype, Year, study.name) %>% 
  dplyr::select(
    study.name,
    Year,
    StudyType,
    subtype,
    PreEyes=PreOpEyes,
    PreIOPMean=PreOpIOPMean,
    PreIOPSD=PreOpIOPStdDev,
    TimeofLastPostOp) 
kable(df_, digits = 1)
```

```{r}
df_ <- df %>% filter(!is.na(LastPeriodAbsIOPChangeMean)) %>% 
  arrange(subtype, Year, study.name) %>% 
  dplyr::select(
    study.name,
    PostEyes=LastPeriodEyes,
    PostIOPMean=LastPeriodIOPMean,
    PostIOPSD=LastPeriodIOPStdDev,
    PostIOPChangeMean=LastPeriodAbsIOPChangeMean,
    PostIOPChangeSD=LastPeriodAbsIOPChangeStdDev) 
kable(df_, digits = 1)
```


# 12 & 15. Risk of bias

A major source of bias is loss of follow-up. Retrospective and prospective 
studies have different potential patterns of lossiness. Most retrospective 
studies reported the IOP before and after surgery *for the same set of eyes*,
which leaves the possibility that some eyes received surgery and were lost in 
follow up, which is very hard to quantify. Prospective studies are less problematic 
in this regard, as the number of eyes lost in follow-up is known.  

In the main text, we present the results of the analysis under a missing 
completely at random (MCAR) assumption, which does do not attempt to correct for 
these biases. 

In a supplementary analysis, we verify the robustness of the results by 
considering the results of including just the prospective studies, which are 
less subject to bias. We also consider how our results would change if the eyes 
lost to follow up showed overall worse outcomes than in the measured eyes. We
consider what would happen if these eyes had 0 change in IOP compared to the
pre-period, or a 5 mm Hg increase compared to the observed eyes. However,
because the loss of follow-up in the prospective group was rather mild (table
below), the results were not very sensitive to this form of loss, except in the
case of PXG, where only two prospective studies have been performed, and they 
both had significant loss in follow up.

```{r message=FALSE}
df <- read.data()
df <- filter.data(df, 'prospective')

df_ <- df %>% filter(subtype != 'acute', !is.na(LastPeriodEyes))
frac <- (df_$PreOpEyes - df_$LastPeriodEyes)/ df_$PreOpEyes
frac.range <- data.frame("fraction with loss"=mean(frac > 0),
           "mean loss"=mean(frac),
           "SD loss"=sd(frac),
           "min loss"=min(frac),
           "max loss"=max(frac), 
           check.names=FALSE)
kable(frac.range, digits=2)
```

Small studies often show larger magnitude effects because of the so-called file 
drawer problem: the tendency of negative results from small-scale studies to 
never get published (ref). We used standard funnel plot based methods to verify 
this was the case. In fact, the data shows an unusual, opposite trend: the
larger studies tended to show the largest effects, an effect we attribute to
washout. We discuss this in detail in the main results.

A final, important source of bias is the lack of a true control group. It would
very difficult to create a true control -- sham surgery -- and unethical to not
subject patients to any treatments. 

> TODO(Marisse): Find the best estimate for the average change in IOP over a
year for patients that have glaucoma and don't change medications. This could be
from a meta-analysis of the change in IOP in placebo arms of drug trials, for
instance. I thought I saw something like that in a paper 

> TODO(Marisse): think of other ways we could address this confound

# 14. Synthesis of results

We used standard random-effects meta-analysis for continuous outcomes (ref 
`meta` package in R; ref book) throughout the text. We used multivariate 
meta-analysis (package `mvmeta`; ref book) to summarize the temporal trend in 
IOP.

# 15. Additional analyses

> TODO(Patrick): explain the washout and medication stuff.
