---
title: "Methods and results"
output: pdf_document
---

Follow the procedures in [PRISMA 2009](http://prisma-statement.org/documents/PRISMA%202009%20checklist.pdf).

[Write about data collection procedures, inclusion criteria, data auditing, 
etc.]

Studies often reported several arms corresponding to different severities, 
treatments and subtypes. For phacoemulsification studies, we aggregated 
the data to obtain one arm per glaucoma subtype (ACG, OAG, PXG, acute) per study. 
For MIGS studies, we aggregated the data to obtain one arm per surgery type 
(Ab interno, canaloplasty, etc.).

After this step, the total number of arms and eyes per subtype was as follows:

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
source('read-data.R')
df <- read.data()
```

```{r kable}
df_ <- df %>% group_by(classification = ifelse(MIGsYorN == 'Y', 'MIGS', as.character(subtype))) %>% 
  dplyr::summarize(
    n.total = n(),
    n.prospective = sum(StudyType == 'Prospective'),
    n.retrospective = sum(StudyType == 'Retrospective'),
    PreOpEyes = sum(PreOpEyes),
    FinalPeriodEyes = sum(LastPeriodEyes))
df_ <- df_ %>% rbind(., summarize(df_, 
                                  classification="(all)",
                                  n.total=sum(n.total),
                                  n.prospective=sum(n.prospective),
                                  n.retrospective=sum(n.retrospective),
                                  PreOpEyes=sum(PreOpEyes),
                                  FinalPeriodEyes = sum(FinalPeriodEyes)))
  
kable(df_)
```

