---
title: "Methods and results"
output:
  pdf_document: 
    latex_engine: xelatex
geometry: margin=0.5in
---

Follow the procedures in [PRISMA
2009](http://prisma-statement.org/documents/PRISMA%202009%20checklist.pdf).

# 10. Data extraction

An undergraduate assistant manually extracted data from the tables of the
identified studies. In three cases where the data of interest was not cited in -
the IOP measurements and medication numbers in (ref Liu et al.), and the 
medication numbers in (ref Hayashi, Tham 2013), the data was extracted from 
figures using a graphical data extraction tool (WebPlotDigitizer,
http://arohatgi.info/WebPlotDigitizer/; ref Cochrane https://www.ncbi.nlm.nih.gov/pubmed/26780258).

The data tables were spot-checked by M.M. against the primary studies. Data was
also audited for self-consistency by P.M. -- e.g. ensuring that differences in
IOP post-surgery were consistent with pre- and post-surgery IOP measurements, 
that number of eyes tracked did not increase through follow-up, etc. The full
extent of these automated checks can be viewed at (ref github 
link).

The data for OAG studies was also cross-verified against Table 1 of Thomas et al
(ref). The full extracted data can be accessed at (ref Github link).

# 11 & 13. Data items & summary measures

We extracted the following columns from the primary sources:

```{r echo=FALSE, warning=FALSE, message=FALSE}
source("read-data.R")
library(knitr)
df <- read.data()
```


```{r message=FALSE}
extracted.data <- read.column.names()
kable(data.frame(extracted.data))
```

> TODO(Marisse): transform this to a table, group similar measurements together.

The main metric of interest for our meta-analysis was the IOP change at the 
final follow-up period (12 months or longer). In complementary analyses, we 
performed meta-analyses separately for 6 month, 12 month, and > 12 month 
follow-ups. When the final period of follow up was not equal for every patient,
we put in the 6-month follow up bin if it was less than 12 months, on average,
and in the > 12 month bin it was more than 12 months.

We also analyzed the change in number of medications between the pre-
period and final follow-up.

## Aggregation and cleaning

Some studies used washout in at least some time periods to measure the effect of
surgery separately from the effect of medication. When both pre- and
post-surgery IOP were reported with washout, we used the washed out IOPs (refs
Pfeiffer, Vold); this is indicated by an asterisk in the main results. When
washout was used in only some periods, we used non-washout values across time
periods, for consistency. The only exception was (ref Azuara−Blanco 2016), where
IOP was measured with washout in the pre-period but not in the post period, due
to study design. We indicate this study with a double asterisk.

Many studies did not report IOP change directly, but rather only absolute IOP in
the pre-period and at the follow-up period. Assuming follow-ups are missing 
completely at random (MCAR), the IOP change can be estimated simply by the difference 
of these measurements, while the SD ($\sigma$) of the IOP change is estimated by:

$\sigma_{drop}^2 = \sigma_{pre}^2 + \sigma_{post}^2 - 2 r \sigma_{pre} \sigma_{post}$

$r$, the correlation between the pre-and-post measurements, was set to the 
median of the correlations in the studies reporting the full set of metrics 
(r = 0.28).

```{r message=FALSE}
df <- read.data(impute.change = FALSE) %>% filter(subtype != 'acute')
rs <- with(df, (PreOpIOPStdDev ** 2 + LastPeriodIOPStdDev ** 2 - LastPeriodAbsIOPChangeStdDev ** 2) / (PreOpIOPStdDev * LastPeriodIOPStdDev)/2)
quantile(rs, .5, na.rm=TRUE)
```

The reported SD of the change in IOP in one study (Jacobi et al. 2002) implied a 
pre-post correlation larger than 1; since the reported range of the IOP change was
also incompatible with the reported range, we replaced this SD with one 
estimated from the reported pre- and post-surgery IOP SDs.

Many studies reported several arms corresponding to different severities, 
treatments and subtypes. For phacoemulsification studies, we aggregated 
the data to obtain one arm per glaucoma subtype (ACG, OAG, PXG, acute) 
per study. 

After this step, the total number of arms and eyes per subtype was as follows:

```{r message=FALSE}
df <- read.data(fill.last=FALSE)
df_ <- df %>% filter(!is.na(LastPeriodAbsIOPChangeMean)) %>% group_by(subtype) %>% 
  dplyr::summarize(
    n.total = n(),
    n.prospective = sum(StudyType == 'Prospective'),
    n.retrospective = sum(StudyType == 'Retrospective'),
    PreOpEyes = sum(PreOpEyes),
    FinalPeriodEyes = sum(LastPeriodEyes))

df_ <- df_ %>% rbind(., summarize(df_, 
                                  subtype="(all)",
                                  n.total=sum(n.total),
                                  n.prospective=sum(n.prospective),
                                  n.retrospective=sum(n.retrospective),
                                  PreOpEyes=sum(PreOpEyes),
                                  FinalPeriodEyes = sum(FinalPeriodEyes)))
kable(df_, caption = "Number of arms with 12 month+ follow up")
```

```{r}
df_ <- df %>% group_by(subtype) %>% 
  dplyr::summarize(
    n.total = n(),
    n.prospective = sum(StudyType == 'Prospective'),
    n.retrospective = sum(StudyType == 'Retrospective'),
    PreOpEyes = sum(PreOpEyes))

df_ <- df_ %>% rbind(., summarize(df_, 
                                  subtype="(all)",
                                  n.total=sum(n.total),
                                  n.prospective=sum(n.prospective),
                                  n.retrospective=sum(n.retrospective),
                                  PreOpEyes=sum(PreOpEyes)))
kable(df_, caption = "Number of arms across all studies")
```

The full list of studies is as follow:

```{r}
df_ <- df %>% filter(!is.na(LastPeriodAbsIOPChangeMean)) %>% 
  arrange(subtype, Year, study.name) %>% 
  dplyr::select(
    study.name,
    Year,
    StudyType,
    subtype,
    PreEyes=PreOpEyes,
    PreIOPMean=PreOpIOPMean,
    PreIOPSD=PreOpIOPStdDev,
    TimeofLastPostOp) 
kable(df_, digits = 1)
```

```{r}
df_ <- df %>% filter(!is.na(LastPeriodAbsIOPChangeMean)) %>% 
  arrange(subtype, Year, study.name) %>% 
  dplyr::select(
    study.name,
    PostEyes=LastPeriodEyes,
    PostIOPMean=LastPeriodIOPMean,
    PostIOPSD=LastPeriodIOPStdDev,
    PostIOPChangeMean=LastPeriodAbsIOPChangeMean,
    PostIOPChangeSD=LastPeriodAbsIOPChangeStdDev) 
kable(df_, digits = 1)
```


# 12 & 15. Risk of bias

A major source of bias is loss of follow-up. Retrospective and prospective 
studies have different potential patterns of lossiness. Most retrospective 
studies reported the IOP before and after surgery *for the same set of eyes*,
which leaves the possibility that some eyes received surgery and were lost in 
follow up, which is very hard to quantify. Prospective studies are less 
problematic in this regard, as the number of eyes lost in follow-up is known.  

In the main text, we present the results of the analysis under a missing 
completely at random (MCAR) assumption, which does do not attempt to correct for 
these biases. 

In a supplementary analysis, we verify the robustness of the results by 
considering the results of including just the prospective studies, which are 
less subject to bias. We also perform a sensitivity analysis, where we estimate
how our results would change under an MNAR (missing not at random) scenario 
where eyes with the worst outcomes are lost to follow up.

Specifically, we consider what would happen if the eyes lost in follow-up had 0
change in IOP compared to the pre-period. We also considered what would happen 
if the change in IOP in the group lost in follow-up was 5 mm Hg higher than in
the observed group (ref Metaanalysis an introduction with R).

Because the loss of follow-up in the prospective group was rather mild (table
below), the results were not very sensitive to this form of loss, except in the
case of PXG, where only two prospective studies have been performed, and they
both had significant loss in follow up.

```{r message=FALSE}
df <- read.data()
df <- filter.data(df, 'prospective')

df_ <- df %>% filter(subtype != 'acute', !is.na(LastPeriodEyes))
frac <- (df_$PreOpEyes - df_$LastPeriodEyes)/ df_$PreOpEyes
frac.range <- data.frame("fraction with loss"=mean(frac > 0),
           "mean loss"=mean(frac),
           "SD loss"=sd(frac),
           "min loss"=min(frac),
           "max loss"=max(frac), 
           check.names=FALSE)
kable(frac.range, digits=2)
```

Small studies often show larger magnitude effects because of the so-called file 
drawer problem: the tendency of negative results from small-scale studies to 
never get published (ref meta-analysis book). We used standard funnel plot based
methods to check if this was the case. In fact, the data showed an unusual,
opposite trend: the larger studies tended to show the largest effects, an effect
we attribute to washout. We discuss this in detail in the main results.

A final, important source of bias is the lack of a true control group. It would
very difficult to create a true control -- sham surgery -- and unethical to not
subject patients to any treatments. We addressed this concern by aggregating 
information from Stage II trials and from prevention studies to get a very coarse
estimate of how large this effect might be (refs) - the estimates range 
from 1 to 2.5 mm Hg. This estimate is necessarily very crude, because the  
test populations are not comparable (ref improper metaanalysis in glaucoma), but 
such an estimate, no matter how coarse, is important. In particular, the estimate
should be taken into account when interpreting the extant data, and also for 
preparing a power analysis in future studies.

# 14. Synthesis of results

We used standard random-effects meta-analysis for continuous outcomes (ref 
`meta` package in R; ref book) throughout the text. 

# 15. Additional analyses

In our main metaanalysis, we estimated the change in IOP that followed 
phacoemulsification. An ideal analysis would instead estimate the *causal* 
effect of surgery. The causal effect of surgery is the change in IOP following
the intervention, holding all other variables fixed, compared to what would have
happened if we did not perform surgery (ref causal impact). In a complementary
analysis, we estimated this causal effect.

Following surgery, glaucoma medications are frequently adjusted; an ideal 
experiment would hold medications fixed. Studies with washout in the pre- and
post- periods - temporary interruption of medication - offer an excellent
approximation to the ideal fixed medication experiment.

For the studies without washout, we estimated a net effect of surgery: the sum
of the measured effect in IOP and the expected change on IOP if medications has
been fixed to their number in the pre-surgery period. Because we did not know
the exact mix of medications used in each study, we used timolol as our
reference. Timolol is the most commonly prescribed glaucoma medication, and in a
recent meta-analysis, it showed an efficacy close to the mean of all tested
medications (3.8 mmHg; ref).

In practice, if a study without washout reported a change in IOP of -2 mmHg, and
a mean decrease in numbers of meds of 1 in the post-period, we attributed a net
decrease of 5.8 mm Hg in IOP. For a study with washout in the pre- and post- 
periods, the attributed net change in IOP was simply the reported change in IOP.

To this net effect, we must subtract the change in IOP that would have occured
in a hypothetical control arm. As outlined in the *Risk of bias* section, we do
not have great estimates of this quantity. In the graph for this analysis, we
indicate our best guess for this quantity (-2 mmHg), based on the extant data.

We believe this is a conservative estimate of the net causal impact of surgery for 
three reasons:

  * Baseline: in studies without washout, the starting baseline was generally
  lower because of the use of drugs to control IOP. It's likely that surgery
  would show a larger effect on IOP when starting from a larger baseline. 
  * Substitution of drugs: studies did not report the mix of medications they
  used on patients. It's possible that more efficacious drugs that were less
  well tolerated were substituted for less powerful drugs with better side
  effect profiles once IOP was relatively well controlled.
  * Intent-to-treat: in the real world, adherence for preventive drugs can be
  low. Surgery does not suffer such an adherence problem, so its net relative
  effect can be amplified compared to meds in an intent-to-treat analysis.
  
# Other things to discuss

  * Very little change across time
  * Sensitivity to loss of follow up
  * Ideal experiment design based on review of literature
  * For PXG, there's high loss of follow up, and few good studies -- more 
    research is necessary.

# Additional references

  * Borenstein et al. (2009), Introduction to Meta-Analysis. [[Link]](https://www.amazon.com/Introduction-Meta-Analysis-Michael-Borenstein/dp/0470057246).
  * Schwarzer et al. (2015), Meta-analysis with R. [[Link]](http://www.springer.com/gp/book/9783319214153).
  * Kadic et al. (2016), Extracting data from figures with software was faster, with higher interrater reliability than manual extraction. [[Link]](https://www.ncbi.nlm.nih.gov/pubmed/26780258).
  * Li & Dickersin (2013), Citation of Previous Meta-analyses on the Same Topic: A Clue to Perpetuation of Incorrect Methods? [[Link]](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4730544/).
  * Li et al. (2016), Comparative Effectiveness of First-Line Medications for Primary Open-Angle Glaucoma: A Systematic Review and Network Meta-analysis. [[Link]](https://www.ncbi.nlm.nih.gov/pubmed/26526633).
  * Imbens & Rubin (2015), Causal Inference for Statistics, Social, and Biomedical Sciences: An Introduction. [[Link]](https://www.amazon.com/Causal-Inference-Statistics-Biomedical-Sciences/dp/0521885884).
