---
title: "Sanity checks"
output: pdf_document
---

# Sanity check data graphically

```{r}
library(ggplot2)
source("read-data.R")
df <- read.data()
```

Check that changes add up.

```{r fig.width=6, fig.height=5}
ggplot(df, aes(x = SixMoIOPMean - PreOpIOPMean, y = SixMoAbsIOPChangeMean, label = study.name)) + 
  geom_point() + 
  coord_cartesian(xlim=c(-10, 0)) + 
  geom_abline() + 
  geom_text_repel()
```
 * Hayashi et al. rejected patients they couldn't follow for 12 months
 * Yudhasompop was retrospective
 * Iancu has full follow up
 * Shoji had low follow up loss
 * Yalvac was retrospective
 
Working model: if there's a delta, and no annotation for number of eyes, it's retrospective.
Sanity check: the delta should be 0

```{r}
df %>% mutate(delta.sixMo = SixMoIOPMean - PreOpIOPMean - SixMoAbsIOPChangeMean) %>% 
  filter(!is.na(delta.sixMo)) %>%
  select(study.name, subtype, delta.sixMo, PreOpEyes, SixMoEyes)
```

Also: the number from Iancy don't match up 100%. Likely source: sloppy math.

Let's see what happens at one year.

```{r}
df %>% mutate(delta.OneY = OneYIOPMean - PreOpIOPMean - OneYAbsIOPChangeMean) %>% 
  filter(!is.na(delta.OneY)) %>%
  select(study.name, subtype, delta.OneY, PreOpEyes, OneYEyes)
```

Problem with assumption: Mathalone doesn't fit this pattern.

 * Mathalone is retrospective
 * The measured IOPs are actually for those in the followed-up group (N = 34)
 * TODO(Patrick): fix the coding.

At the last point in the study:

```{r}
df %>% mutate(delta.LastPeriod = LastPeriodIOPMean - PreOpIOPMean - LastPeriodAbsIOPChangeMean) %>% 
  filter(!is.na(delta.LastPeriod)) %>%
  select(study.name, subtype, delta.LastPeriod, PreOpEyes, LastPeriodEyes)
```

That fits into this general pattern. Now let's look at those with no delta (only absolutes).

```{r}
df %>% filter(is.na(SixMoAbsIOPChangeMean) & !is.na(SixMoIOPMean)) %>%
  select(study.name, subtype, PreOpIOPMean, SixMoIOPMean, PreOpEyes, SixMoEyes)
```

 * Euswas is retrospective
 * Lam had full follow up except one eye that died after 17 months (lung cancer)
 * Pachimkul was prospective and had full follow up

```{r fig.width=6, fig.height=5}
ggplot(df, aes(x = OneYIOPMean - PreOpIOPMean, y = OneYAbsIOPChangeMean, label = study.name)) + 
  geom_point() + 
  coord_cartesian(xlim=c(-10, 0)) + 
  geom_abline() + 
  geom_text_repel()
```

```{r fig.width=6, fig.height=5}
ggplot(df, aes(x = LastPeriodIOPMean - PreOpIOPMean, y = LastPeriodAbsIOPChangeMean, label = study.name)) + 
  geom_point() + 
  coord_cartesian(xlim=c(-25, 0)) + 
  geom_abline() + 
  geom_text_repel()
```


Check the relationship between RxPreOpMean and s.d.
```{r}
ggplot(df, aes(x = RxPreOpMean, y = RxPreOpStdDev)) + 
  geom_point() + 
  coord_cartesian(y=c(0, 1.5)) + 
  stat_function(fun = function(x) sqrt((x - floor(x)) * (1 - (x - floor(x)))), color="gray70")
```

```{r}
ggplot(df, aes(x = RxPostOpMean, y = RxPostOpStdDev)) + geom_point() + 
  stat_function(fun = function(x) sqrt((x - floor(x)) * (1 - (x - floor(x)))), color="gray70")
```

Examine loss at one year.

```{r fig.width=6, fig.height=5}
ggplot(df, aes(x=PreOpEyes, y=OneYEyes, label=study.name)) + 
  geom_point(color="red") + 
  geom_abline() + 
  geom_text_repel() + coord_cartesian(xlim=c(0, 250), ylim=c(0, 250))
```

```{r fig.width=6, fig.height=5}
ggplot(df, aes(x=OneYEyes / PreOpEyes, label=study.name)) + 
  geom_histogram() + 
  coord_cartesian()
```

```{r}
ggplot(df, aes(x=subtype, fill=MIGsYorN)) + geom_histogram(stat="count", position = 'dodge')
```

Look at the distribution of eyes and IOP means. 

```{r}
ggplot(df, aes(x=PreOpEyes, y=PreOpIOPMean)) + geom_point()
```

Look at number of eyes and standard deviation.

```{r fig.width=11, fig.height=11}
ggplot(df, aes(x=PreOpEyes, y=PreOpIOPStdDev, label=study.name)) + geom_point(color="red") + geom_text_repel(nudge_x = 5)
```

```{r fig.width=5, fig.height=5}
ggplot(df, aes(x=PreOpEyes, y=SixMoAbsIOPChangeStdDev, label=study.name)) + geom_point(color="red") + geom_text_repel(nudge_x = 5)
```

```{r fig.width=5, fig.height=5}
ggplot(df, aes(x=PreOpEyes, y=SixMoIOPStdDev, label=study.name)) + geom_point(color="red") + geom_text_repel(nudge_x = 5)
```

```{r fig.width=5, fig.height=5}
ggplot(df, aes(x=PreOpEyes, y=OneYAbsIOPChangeStdDev, label=study.name)) + geom_point(color="red") + geom_text_repel(nudge_x = 5)
```

```{r fig.width=5, fig.height=5}
ggplot(df, aes(x=PreOpEyes, y=OneYIOPStdDev, label=study.name)) + geom_point(color="red") + geom_text_repel(nudge_x = 5)
```

```{r fig.width=5, fig.height=5}
ggplot(df, aes(x=PreOpEyes, y=LastPeriodAbsIOPChangeStdDev, label=study.name)) + geom_point(color="red") + geom_text_repel(nudge_x = 5)
```

```{r fig.width=5, fig.height=5}
ggplot(df, aes(x=PreOpEyes, y=SixMoIOPStdDev, label=study.name)) + geom_point(color="red") + geom_text_repel(nudge_x = 5)
```
